---
title: "Exploratory Data Analysis"
author: "ERICK_GURU"
date: "2024-03-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis (EDA)

Exploratory Data Analysis is one of the critical processes of performing initial investigations on data analysis.

Basic idea is to discover the patterns, anomalies, test hypotheses, and check the assumptions with the help of summary statistics and graphical representations.

The main idea about exploratory data analysis are

1.  maximize insight into a data set;

2.  uncover underlying structure;

3.  extract important variables;

4.  detect outliers and anomalies;

5.  test underlying assumptions;

6.  develop parsimonious models;

7.  determine optimal factor settings.

```{r message=FALSE, warning=FALSE}
#libraries
library(tidyverse)
library(DataExplorer)
```

The dataset we are using for exploratory data analysis is from gss_cat.

```{r warning=FALSE, message=FALSE}
gss_cat %>% dim()
head(gss_cat)
tail(gss_cat)
```

The datasets contains 21,483 observations and 9 variables.

### Dataset Overview

```{r}
gss_cat %>%  glimpse()
```

Above function will provide basic idea about variables in the datasets.

### Introduce

Let's introduce the data set based on introduce function.

```{r}
gss_cat %>%  introduce()
```

Now you can see the dataset contains 21483 observations with 9 variables all other kind of information's.

### Visualization

Most EDA techniques are graphical in nature with a few quantitative techniques. Here we are going to explain different kind of visualization techniques to identify the pattern and relationships.

```{r}
gss_cat %>% plot_intro()
```

66.7% contains Discrete columns, 33.3% contains continuous columns.

52.6% contains complete rows and 5.3% contains missing observations.

All columns with missing values are 0%.

```{r}
gss_cat %>% plot_missing()
```

```{r}
gss_cat %>% profile_missing()
```

### Density Plot

```{r}
gss_cat %>%   plot_density()
```

### Histogram

```{r}
gss_cat %>% plot_histogram()
```

### Bar Plot

```{r}
gss_cat %>% plot_bar()
```

### Correlation Plot

```{r}
gss_cat %>% plot_correlation()
```

```{r}
gss_cat %>% plot_correlation(maxcat =6)
```

maxcat can adjust based on your variables and interest. You can make use Data Explorer package for quick checking before proceeding to further analysis.

```{r}
# gss_cat %>%
#  create_report()
# # output_file = "gss_survey_data_profile_report",
# # output_dir = "D:/RStudio/EDA/",
# # y="rincome",
# # report_title = "EDA Report"
# # )
```
